### 2022-07-19 article.tex
Add Greek unicode letters

% This is a template for documentclass{article}.
% See https://ctan.org/ for available packages and documentation.

% this is the beginning of the preamble section
% must always be first line of a <path.tex> document
\documentclass[10pt, letterpaper, twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[sfdefault]{noto}
\setlength{\textwidth}{6in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-1in}
\setlength{\oddsidemargin}{.25in}
\setlength{\evensidemargin}{.25in}

% this package has better control of title, author elements
\usepackage{titling}
\setlength{\droptitle}{-8pt}
\pretitle{\begin{flushleft}\LARGE\bfseries}
\posttitle{\par\end{flushleft}}
\preauthor{\begin{flushleft}\normalsize}
\postauthor{\end{flushleft}}
\predate{\begin{flushleft}\small}
\postdate{\end{flushleft}}
\date{\today}
% edit the following line for the title of the article
\title{article title}
% edit the following line for the author(s) of the article
\author{article author}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\raggedbottom
\renewenvironment{abstract}{\par\noindent\textbf{\abstractname}\par}

% this package has better control over the date format
\usepackage{datetime2}

% this package has better control over 'floats'
\usepackage{float}

% this package has better control over images
\usepackage{graphicx}
\graphicspath{{fot/}}
\usepackage[export]{adjustbox}
\usepackage[singlelinecheck=false,justification=justified]{caption}

% very useful for mathematics
\usepackage{amsmath, amsthm, amssymb}

% very useful for hyperlinks
\usepackage[unicode]{hyperref}
\hypersetup{
    colorlinks=true,
    urlcolor=blue
}

% useful for Greek letters in text mode
\usepackage[cbgreek]{textgreek}

% required for bibtex URLs
\usepackage{url}

% to create a list of equations
\usepackage{tocloft}
\setlength{\cfttabindent}{0in}
\setlength{\cftfigindent}{0in}

% used to create tables
\usepackage{booktabs}

% useful for the bibliography
\usepackage[style=numeric]{biblatex}
\addbibresource{pilon_gilles.bib}

% define theorem-like elements
\theoremstyle{myenv}
\theoremstyle{mytheorem}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{axiom}[theorem]{Axiom}
\theoremstyle{mydefinition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{question}[theorem]{Question}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\makeatletter
\@addtoreset{theorem}{section}
\makeatother

% set no indent for lists
\usepackage{enumitem}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{leftmargin=*}

% create a list of equations command
\newcommand{\listequationsname}{List of Equations}
\newlistof{myequations}{equ}{\listequationsname}
\newcommand{\myequations}[1]{%
\addcontentsline{equ}{myequations}{\protect\numberline{\theequation}#1}\par}
\renewcommand{\cftequtitlefont}{\normalfont\Large\bfseries}

% create a list of Greek letters command
\newcommand{\listgreekname}{List of Greek Letters}
\newlistof{mygreek}{grk}{\listgreekname}
\newcommand{\mygreek}[1]{%
\addcontentsline{grk}{mygreek}{\protect\numberline{\thegreek}#1}\par}
\renewcommand{\cftgrktitlefont}{\normalfont\Large\bfseries}

% create a list of theorems command
\newcommand{\listtheoremsname}{List of Theorems}
\newlistof{mytheorems}{thm}{\listtheoremsname}
\newcommand{\mytheorems}[1]{%
\addcontentsline{thm}{mytheorems}{\protect\numberline{\thetheorem}#1}\par}
\renewcommand{\cftthmtitlefont}{\normalfont\Large\bfseries}

% create a list of definitions command
\newcommand{\listdefinitionsname}{List of Definitions}
\newlistof{mydefinitions}{def}{\listdefinitionsname}
\newcommand{\mydefinitions}[1]{%
\addcontentsline{def}{mydefinitions}{\protect\numberline{\thedefinition}#1}\par}
\renewcommand{\cftdeftitlefont}{\normalfont\Large\bfseries}

% create a list of corollaries command
\newcommand{\listcorollariesname}{List of Corollaries}
\newlistof{mycorollaries}{cor}{\listcorollariesname}
\newcommand{\mycorollaries}[1]{%
\addcontentsline{cor}{mycorollaries}{\protect\numberline{\thecorollary}#1}\par}
\renewcommand{\cftcortitlefont}{\normalfont\Large\bfseries}

% create a list of lemmas command
\newcommand{\listlemmasname}{List of Lemmas}
\newlistof{mylemmas}{lem}{\listlemmasname}
\newcommand{\mylemmas}[1]{%
\addcontentsline{lem}{mylemmas}{\protect\numberline{\thelemma}#1}\par}
\renewcommand{\cftlemtitlefont}{\normalfont\Large\bfseries}

% create a list of propositions command
\newcommand{\listpropositionsname}{List of Propositions}
\newlistof{mypropositions}{prop}{\listpropositionsname}
\newcommand{\mypropositions}[1]{%
\addcontentsline{prop}{mypropositions}{\protect\numberline{\theproposition}#1}\par}
\renewcommand{\cftproptitlefont}{\normalfont\Large\bfseries}

% useful for epigraphs
\usepackage{epigraph}

% this is the end of the preamble section

\begin{document}
\maketitle

\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
\end{abstract}

\newpage
\tableofcontents

\newpage
\section{Introduction}
\label{sec:introduction}
\renewcommand{\epigraphflush}{flushleft}
\renewcommand{\textflush}{flushleft}
\renewcommand{\sourceflush}{flushleft}
\epigraph{Remember, the objective is to take the right action on the process. No
credit is given for finding the `right numbers.'}{Donald J. Wheeler}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{Prior research}
\label{sec:prior_research}
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{Results}
\label{sec:results}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\newpage
\section{Bold, italics, and underline}
\label{sec:bold_italics_underline}
Some of the \textbf{greatest} \textit{discoveries} in \underline{science} were made by \textbf{\textit{accident}}.

\section{URLs}
\label{sec:urls}
Hyperlinks within the document can be easily added. For example, see Graphics \ref{sec:graphics} and Anatomy of a matplotlib figure \ref{fig:anatomy_of_a_matplotlib_figure} and Definition of a limit \ref{def:limit}.

Hyperlinks to external URLs can be easily added. For example, see \href{https://en.wikipedia.org/wiki/Latex}{Wikipedia \LaTeX}.
\section{Lists}
\subsection{Unordered list}
\label{sec:unordered_list}
\begin{itemize}
    \item Illinois
        \begin{itemize} \item Chicago \item Naperville \end{itemize}
    \item Kentucky
        \begin{itemize} \item Lexington \item Louisville \end{itemize}
\end{itemize}

\subsection{Ordered list}
\label{sec:ordered_list}
\begin{enumerate}
    \item Cat
        \begin{enumerate}
            \item Persian
            \item Siamese
        \end{enumerate}
    \item Dog
        \begin{enumerate}
            \item Aussiedoodle
            \item Newfoundland
        \end{enumerate}
\end{enumerate}

\newpage
\section{Graphics}
\label{sec:graphics}
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

% H means 'here definitely' to place the figure
\begin{figure}[H]
% \includegraphics[width=\textwidth, inner]{matplotlib_anatomy_of_a_figure}
\includegraphics[width=3in, left]{matplotlib_anatomy_of_a_figure}
\caption{Anatomy of a matplotlib figure}
\end{figure}
\label{fig:anatomy_of_a_matplotlib_figure}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\newpage
\section{Tables}
\label{sec:tables}

\begin{table}[H]
\begin{flushleft}
\begin{tabular}{lcccccc}
\toprule
& I &  II & III & IV & V & VI \\
\midrule
Vandali     & 123 & 456 & 678 & 321 & 644 & 768  \\
Visigothorum & 021 & 229 & 678 & 123 & 456 & 678 \\
\bottomrule
\end{tabular}
\label{tab:visigothi_cum_romanis}
\caption{Visigothi cum Romanis}
\end{flushleft}
\end{table}

\begin{table}[H]
\begin{flushleft}
\begin{tabular}{ccc}
\toprule
$A$ & $B$ & $ A \land B$ \\
\midrule
T & T & T \\
T & F & F \\
F & T & F \\
F & F & F \\
\bottomrule
\end{tabular}
\label{tab:boolean_truth_table}
\caption{Boolean truth table}
\end{flushleft}
\end{table}

\newpage
\section{Mathematics}
\label{sec:mathematics}

\subsection{Inline mathematics}
\label{sec:inline_mathematics}`
Examples of inline mathematics $\frac{a}{b}$, followed by text and another inline equation $\lim_{x\to 0}\frac{\sin x}{x}$=1 and more text $\sum_{k=0}^\infty \frac{1}{2^k}=2$ and finally much more text to cover many lines of a paragraph.

\subsection{Display-style mathematics}
\label{sec:display_style_mathematics}
Examples of mathematics on their own lines. All equations are flush-left to the page unless multilines, and thus aligned on a symbol with the longest equation flush-left. The equations numbers are flush-right.

\begin{flalign}
    \text{measured observation} = \text{true value} + \text{measurement system error}&&
\end{flalign}
\myequations{Measured observation}

\begin{flalign}
    \text{ICC} = \frac{\text{variance of the product values}}{\text{variance of the product observations}}&&
\end{flalign}
\myequations{Intraclass correlation coefficient}

\begin{flalign}
    \overline{X}  \pm \text{CI} & = \frac{\sum\limits_{i=1}^{n} X_i}{n} \pm t_{1- \alpha / 2 \, \text{,} \, n-1} \times \frac{s}{\sqrt{n}}&&
\end{flalign}
\myequations{Average $\pm$ confidence interval}

\begin{flalign}
    \text{distance to target} & = \text{specification target} - \text{location statistic} = T - \overline{X}&&
\end{flalign}
\myequations{Distance to target}

\begin{flalign}
    s \pm \text{CI} & = \sqrt{\frac{\sum\limits_{i=1}^{n} \left(X_i - \overline{X}\right)^2}{n - 1}} \pm \left(\sqrt{\frac{(n - 1)\,s^{2}}{\chi_{\alpha / 2}^2}},\sqrt{\frac{(n - 1)\,s^{2}}{\chi_{1 - {\alpha / 2}}^2}}\right)&&
\end{flalign}
\myequations{Standard deviation $\pm$ confidence interval}

\begin{flalign}
    \text{distance to nearest specification} & = \frac{\text{nearest specification} - \text{location statistic}}{\text{dispersion statistic}}&& \\
                                             & = \frac{\text{min}\,(\overline{X} - \text{LSL}\,, \text{USL} - \overline{X})}{s}&&
\end{flalign}
\myequations{Distance to nearest specification}

The confidence intervals are:

\begin{flalign}
    \text{Lower bound} = \hat{P}_{pk} - z_{1 - \alpha / 2} \sqrt{\frac{1}{\left(\frac{\text{toler}}{2}\right)^2 n} + \frac{\left(\hat{P}_{pk}\right)^2}{2 \nu}}&&
\end{flalign}
\myequations{$\hat{P}_{pk}$ lower bound}

\begin{flalign}
    \text{Upper bound} = \hat{P}_{pk} + z_{1- \alpha / 2} \sqrt{\frac{1}{\left(\frac{\text{toler}}{2}\right)^2 n} + \frac{\left(\hat{P}_{pk}\right)^2}{2 \nu}}&&
\end{flalign}
\myequations{$\hat{P}_{pk}$ upper bound}

where:

\begin{flalign*}
    n                           & = \text{the number of observations in the sample}&& \\
    \nu                         & = \text{the degrees of freedom $= n - 1$}&& \\
    \chi_{\alpha \text{,}\nu}^2 & = \text{the $\alpha$ percentile of the chi-square distribution with $\nu$ degrees of freedom}&& \\
    \text{toler}                & = \text{multiplier of the standard deviation tolerance (Minitab\textregistered\ uses 6 as the default value)}&& \\
    z_{1 - \alpha / 2}          & = \text{the $1 - \alpha / 2$ percentile of the standard normal distribution}&& \\
\end{flalign*}

\begin{flalign}
A=\begin{pmatrix}
1 &0 &4 &-1 \\
0 &1 &-3 &1 \\
0 &0 &0 &1
\end{pmatrix}&&
\end{flalign}

\begin{flalign}
A = \begin{bmatrix}
1 &0 &4 &-1\\
0 &1 &-3 &1 \\
0 &0 &0 &1
\end{bmatrix}&&
\end{flalign}

\newpage
\section{Greek letters}
\label{sec:greek_letters}

\subsection{Greek unicode}
\label{sec:greek_unicode}

Greek letters can be inserted in mathematics and text using Unicode characters. In Neovim, the sequence is `ControlKey + v uhhhh` where `hhhh` is the Unicode.

alpha 03b1 α 0391 Α \\
beta 03b2 β 0392 Β \\
gamma 03b3 γ 0393 Γ \\
delta 03b4 δ 0394 Δ \\
epsilon 03b5 ε 0395 Ε \\
zeta 03b6 ζ 0396 Ζ \\
eta 03b7 η 0397 Η \\
theta 03b8 θ 0398 Θ \\
iota 03b9 ι 0399 Ι \\
kappa 03b1 α 039a Κ \\
lambda 03bb λ 039b Λ \\
mu 03bc μ 039c Μ \\
nu 03bd ν 039d Ν \\
xi 03be ξ 039e Ξ \\
omicron o O \\
pi 03c0 π 03ao Π \\
rho 03c1 ρ 03a1 Ρ \\
sigma 03c3 σ 03a3 Σ \\
tau 03c4 τ 03a4 Τ \\
upsilon 03c5 υ 03a5 Υ \\
phi 03c6 φ 03a6 Φ \\
chi 03c7 χ 03a7 Χ \\
psi 03c8 ψ 03a8 Ψ \\
omega 03c9 ω 03a9 Ω \\

\subsection{Greek mathematics}
\label{sec:greek_mathematics}

alpha $\alpha$ A \\
beta $\beta$ B \\
gamma $\gamma$ $\Gamma$ \\
delta $\delta$ $\Delta$ \\
epsilon$\epsilon$ E \\
zeta $\zeta$ Z \\
eta $\eta$ H\\
theta $\theta$ $\Theta$ \\
iota $\iota$ I \\
kappa $\kappa$ K \\
lambda $\lambda$ $\Lambda$ \\
mu $\mu$ M \\
nu $\nu$ N \\
xi $\xi$ $\Xi$ \\
omicron o O \\
pi $\pi$ $\Pi$\\
rho $\rho$ P \\
sgima $\sigma$ $\Sigma$ \\
tau $\tau$ T \\
upsilon $\upsilon$ $\Upsilon$ \\
phi $\phi$ $\Phi$ \\
chi $\chi$ X \\
psi $\psi$ $\Psi$ \\
omega $\omega$ $\Omega$ \\

\subsection{Greek text}
\label{sec:greek_text}

alpha \textalpha \textAlpha \\
beta \textbeta \textBeta \\
gamma \textgamma \textGamma \\
delta \textdelta \textDelta \\
epsilon \textepsilon \textEpsilon \\
zeta \textzeta \textZeta \\
eta \texteta \textEta \\
theta \texttheta \textTheta \\
iota \textiota \textIota \\
kappa \textkappa \textKappa \\
lambda \textlambda \textLambda \\
mu \textmu \textMu \\
nu \textnu \textNu \\
xi \textxi \textXi \\
omicron o O \\
pi \textpi \textPi \\
rho \textrho \textRho \\
sigma \textsigma \textSigma \\
tau \texttau \textTau \\
upsilon \textupsilon \textUpsilon \\
phi \textphi \textPhi \\
chi \textchi \textChi \\
psi \textpsi \textPsi \\
omega \textomega \textOmega \\

\newpage
\section{Theorems}
\label{sec:theorems}

\begin{theorem}[Fundamental Theorem of Calculus]
\label{thm:ftc}
If $f$ is continuous on the interval $[a,b]$, then \[\frac{d}{dx}\int_a^x f(t)dx=f(x)\] for all $a\le x\le b$.
\end{theorem}
\mytheorems{Fundamental theorem of calculus}

\section{Definitions}
\label{sec:definitions}

\begin{definition}[Definition of limit]
\label{def:limit}
Let $f$ be a function defined on an open interval containing $x=a$, but perhaps not at $x=a$.  We say $\displaystyle\lim_{x\to a}f(x)=L$ if for any $\epsilon>0$ there exists $\delta>0$ such that $|f(x)-L|<\epsilon$ whenever $0<|x-a|<\delta$.
\end{definition}
\mydefinitions{Definition of limit}

\begin{definition}
\label{def:derivative}
The \textit{derivative} of $f$ is defined to be $\displaystyle f^\prime(x)=\lim_{h\to 0}\frac{f(x+h)-f(x)}{h}$, if this limit exists.
\end{definition}
\mydefinitions{Derivative of f}

\section{Corollaries}
\label{sec:corollaries}
\begin{corollary}
If $f$ is continuous on the interval $[a,b]$, then \[\int_a^b f(x)dx=F(b)-F(a)\] where $F$ is any antiderivative of $f$.
\end{corollary}
\mycorollaries{f is continuous}

\newpage
\section{Lemmas}

\label{sec:lemmas}
\begin{lemma}
$x+y=z$
\end{lemma}
\mylemmas{Lem 1}

\begin{lemma}
$z-r+t=3$
\end{lemma}
\mylemmas{Lem 2}

\begin{lemma}
$q=r$
\end{lemma}
\mylemmas{Lem 3}

\begin{lemma}
$x+y=z$
\end{lemma}
\mylemmas{Lem 4}

\begin{lemma}
$z-r+t=3$
\end{lemma}
\mylemmas{Lem 5}

\begin{lemma}
$q=r$
\end{lemma}
\mylemmas{Lem 6}

\newpage
\section{Propositions}

\label{sec:propositions}
\begin{proposition}
$x+y=z$
\end{proposition}
\mypropositions{Prop 1}

\begin{proposition}
$\alpha=1-x$
\end{proposition}
\mypropositions{Prop 2}

\newpage
\section{Citations}
\label{sec:citations}

% The definitive book on MSA is by Wheeler \cite{wheeler2006}. The definitive book on SPC is by Wheeler \cite{wheeler2004}. The original article on Taguchi indices was by Russell Boyles \cite{boyles1991}. A very good article on the $C_\text{pm}$ index is found in Chen et.al. \cite{chan1988}. Gunter \cite{gunter1989}.

Here is a single citation without prenote and postnote. The definitive book on MSA is by Wheeler \cite{wheeler2006}.

Here is a single citation with prenote only. The definitive book on SPC is by \cite[Wheeler][]{wheeler2006}.

Here is a single citation with prenote and postnote. For further details on $C_\text{pm}$, \cite[see][pages 1-10]{chan1988}.

\newpage
\section{To do}

\begin{itemize}
    \item list to do items here
\end{itemize}

\newpage
\listoffigures

\newpage
\listoftables

\newpage
\listofmyequations

\newpage
\listofmytheorems

\newpage
\listofmydefinitions

\newpage
\listofmycorollaries

\newpage
\listofmylemmas

\newpage
\listofmypropositions

\newpage
\section{Acknowledgements}
\label{sec:acknowledgements}

The author wishes to thank ...

\section{Supporting information}
\label{sec:supporting_information}

Add files, data, and other supporting information, and where it can be obtained.

\newpage
\section{References}
\nocite{*}
\printbibliography[heading=none]
\label{sec:references}

\end{document}