### 2022-07-20 article.tex
Tidy section, subsection lables

% This is a template for documentclass{article}.
% See https://ctan.org/ for available packages and documentation.

% this is the beginning of the preamble section
% must always be first line of a <path.tex> document
\documentclass[10pt, letterpaper, twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[sfdefault]{noto}
\setlength{\textwidth}{6in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-1in}
\setlength{\oddsidemargin}{.25in}
\setlength{\evensidemargin}{.25in}

% this package has better control of title, author elements
\usepackage{titling}
\setlength{\droptitle}{-8pt}
\pretitle{\begin{flushleft}\LARGE\bfseries}
\posttitle{\par\end{flushleft}}
\preauthor{\begin{flushleft}\normalsize}
\postauthor{\end{flushleft}}
\predate{\begin{flushleft}\small}
\postdate{\end{flushleft}}
\date{\today}
% edit the following line for the title of the article
\title{article title}
% edit the following line for the author(s) of the article
\author{article author}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\raggedbottom
\renewenvironment{abstract}{\par\noindent\textbf{\abstractname}\par}

% this package has better control over the date format
\usepackage{datetime2}

% this package has better control over 'floats'
\usepackage{float}

% this package has better control over images
\usepackage{graphicx}
\graphicspath{{fot/}}
\usepackage[export]{adjustbox}
\usepackage[singlelinecheck=false,justification=justified]{caption}

% very useful for mathematics
\usepackage{amsmath, amsthm, amssymb}

% very useful for hyperlinks
% useful for Greek letters in text mode
\usepackage[cbgreek]{textgreek}

% to create a list of equations
\usepackage{tocloft}
\setlength{\cfttabindent}{0in}
\setlength{\cftfigindent}{0in}

% used to create tables
\usepackage{booktabs}
\usepackage{multirow}

% useful for the bibliography
\usepackage[style=authoryear]{biblatex}
\addbibresource{pilon_gilles.bib}

% define theorem-like elements
\theoremstyle{myenv}
\theoremstyle{mytheorem}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{axiom}[theorem]{Axiom}
\theoremstyle{mydefinition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{question}[theorem]{Question}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\makeatletter
\@addtoreset{theorem}{section}
\makeatother

% set no indent for lists
\usepackage{enumitem}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{leftmargin=*}

% create a list of equations command
\newcommand{\listequationsname}{List of Equations}
\newlistof{myequations}{equ}{\listequationsname}
\newcommand{\myequations}[1]{%
\addcontentsline{equ}{myequations}{\protect\numberline{\theequation}#1}\par}
\renewcommand{\cftequtitlefont}{\normalfont\Large\bfseries}

% create a list of Greek letters command
\newcommand{\listgreekname}{List of Greek Letters}
\newlistof{mygreek}{grk}{\listgreekname}
\newcommand{\mygreek}[1]{%
\addcontentsline{grk}{mygreek}{\protect\numberline{\thegreek}#1}\par}
\renewcommand{\cftgrktitlefont}{\normalfont\Large\bfseries}

% create a list of theorems command
\newcommand{\listtheoremsname}{List of Theorems}
\newlistof{mytheorems}{thm}{\listtheoremsname}
\newcommand{\mytheorems}[1]{%
\addcontentsline{thm}{mytheorems}{\protect\numberline{\thetheorem}#1}\par}
\renewcommand{\cftthmtitlefont}{\normalfont\Large\bfseries}

% create a list of definitions command
\newcommand{\listdefinitionsname}{List of Definitions}
\newlistof{mydefinitions}{def}{\listdefinitionsname}
\newcommand{\mydefinitions}[1]{%
\addcontentsline{def}{mydefinitions}{\protect\numberline{\thedefinition}#1}\par}
\renewcommand{\cftdeftitlefont}{\normalfont\Large\bfseries}

% create a list of corollaries command
\newcommand{\listcorollariesname}{List of Corollaries}
\newlistof{mycorollaries}{cor}{\listcorollariesname}
\newcommand{\mycorollaries}[1]{%
\addcontentsline{cor}{mycorollaries}{\protect\numberline{\thecorollary}#1}\par}
\renewcommand{\cftcortitlefont}{\normalfont\Large\bfseries}

% create a list of lemmas command
\newcommand{\listlemmasname}{List of Lemmas}
\newlistof{mylemmas}{lem}{\listlemmasname}
\newcommand{\mylemmas}[1]{%
\addcontentsline{lem}{mylemmas}{\protect\numberline{\thelemma}#1}\par}
\renewcommand{\cftlemtitlefont}{\normalfont\Large\bfseries}

% create a list of propositions command
\newcommand{\listpropositionsname}{List of Propositions}
\newlistof{mypropositions}{prop}{\listpropositionsname}
\newcommand{\mypropositions}[1]{%
\addcontentsline{prop}{mypropositions}{\protect\numberline{\theproposition}#1}\par}
\renewcommand{\cftproptitlefont}{\normalfont\Large\bfseries}

% useful for epigraphs
\usepackage{epigraph}

\usepackage[unicode]{hyperref}
\hypersetup{
    colorlinks=true,
    urlcolor=magenta,
    citecolor=green,
    linkcolor=red
}

% this is the end of the preamble section

\begin{document}
\maketitle

\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
\end{abstract}

\newpage
\tableofcontents

\newpage
\section{Introduction}\label{sec:introduction}
\renewcommand{\epigraphflush}{flushleft}
\renewcommand{\textflush}{flushleft}
\renewcommand{\sourceflush}{flushleft}
\epigraph{Remember, the objective is to take the right action on the process. No
credit is given for finding the `right numbers.'}{Donald J. Wheeler}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{Prior research}\label{sec:prior_research}
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{Results}\label{sec:results}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\newpage
\section{Bold, italics, and underline}\label{sec:bold_italics_underline}
Some of the \textbf{greatest} \textit{discoveries} in \underline{science} were made by \textbf{\textit{accident}}.

\section{URLs}\label{sec:urls}
Hyperlinks within the document can be easily added. For example, \hyperref[sec:graphics]{see Graphics} and \hyperref[fig:anatomy_of_a_matplotlib_figure]{Anatomy of a matplotlib figure} and \hyperref[def:limit]{Definition of a limit}.

Hyperlinks to external URLs can be easily added. For example, see \href{https://en.wikipedia.org/wiki/Latex}{Wikipedia \LaTeX}.
\section{Lists}\label{sec:lists}
\subsection{Unordered list}\label{sec:unordered_list}
\begin{itemize}
    \item Illinois
        \begin{itemize} \item Chicago \item Naperville \end{itemize}
    \item Kentucky
        \begin{itemize} \item Lexington \item Louisville \end{itemize}
\end{itemize}

\subsection{Ordered list}\label{sec:ordered_list}
\begin{enumerate}
    \item Cat
        \begin{enumerate}
            \item Persian
            \item Siamese
        \end{enumerate}
    \item Dog
        \begin{enumerate}
            \item Aussiedoodle
            \item Newfoundland
        \end{enumerate}
\end{enumerate}

\newpage
\section{Graphics}\label{sec:graphics}
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

% H means 'here definitely' to place the figure
\begin{figure}[H]
% \includegraphics[width=\textwidth, inner]{matplotlib_anatomy_of_a_figure}
\includegraphics[width=3in, left]{matplotlib_anatomy_of_a_figure}
\caption{Anatomy of a matplotlib figure}
\end{figure}
\label{fig:anatomy_of_a_matplotlib_figure}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\newpage
\section{Tables}\label{sec:tables}

\subsection{Simple table}\label{sec:simple_table}`

\begin{table}[H]
\begin{flushleft}
\begin{tabular}{lcccccc}
\toprule
             & I   & II  & III & IV  & V   & VI \\
\midrule
Vandali      & 123 & 456 & 678 & 321 & 644 & 768  \\
Visigothorum & 021 & 229 & 678 & 123 & 456 & 678 \\
\bottomrule
\end{tabular}
\label{tab:visigothi_cum_romanis}
\caption{Visigothi cum Romanis}
\end{flushleft}
\end{table}

\subsection{Multirow table}\label{sec:multirow_table}`
\begin{table}[H]
\footnotesize
\centering
\begin{tabular}{llll}
\toprule
                                           & \textbf{Process requirement}            & \textbf{Sampling}                        & \textbf{Comments} \\
\midrule
\multirow[t]{4}{*}{$C_\text{pk}$}          & - Stable and predictable variation      & - Rational                               & - Possible cost to increasing $C_\text{pk}$ \\
                                           & - Normal distribution                   & - Representative of process              & - Wide confidence interval \\
                                           & - Need at least one specification limit & - Ensure same process conditions         & \\
                                           &                                         & - No identifiable condition changes      & \\
\multirow[t]{4}{*}{$P_\text{pk}$}          & - Stable and predictable variation      & - Random                                 & - Possible cost to increasing ppk \\
                                           & - Normal distribution                   & - Sufficiently large, representative     & - Wide confidence interval \\
                                           & - Need at least one specification limit & - Ensure same process conditions         & - Easiest formulae to explain \\
                                           &                                         & - No identifiable condition changes      & \\
\multirow[t]{4}{*}{$C_\text{pm}$}          & - Stable and predictable variation      & - Same as $C_\text{pk}$ or $P_\text{pk}$ & - $C_\text{pm}$ penalizes for being off-target \\
                                           & - Normal distribution                   &                                          & - Wide confidence interval \\
                                           & - Need at least one specification limit &                                          & - Most difficult formulae to explain \\
                                           & - Need specification target             &                                          & \\
\bottomrule
\end{tabular}
\caption{Comparison of capability indices}
\label{tab:comparison_capability_indices}
\end{table}

\newpage
\section{Mathematics}\label{sec:mathematics}

\subsection{Inline mathematics}\label{sec:inline_mathematics}`
Examples of inline mathematics $\frac{a}{b}$, followed by text and another inline equation $\lim_{x\to 0}\frac{\sin x}{x}$=1 and more text $\sum_{k=0}^\infty \frac{1}{2^k}=2$ and finally much more text to cover many lines of a paragraph.

\subsection{Display-style mathematics}\label{sec:display_style_mathematics}
Examples of mathematics on their own lines. All equations are flush-left to the page unless multilines, and thus aligned on a symbol with the longest equation flush-left. The equations numbers are flush-right.

\begin{flalign}
    \text{measured observation} = \text{true value} + \text{measurement system error}&&
\end{flalign}
\myequations{Measured observation}

\begin{flalign}
    \text{ICC} = \frac{\text{variance of the product values}}{\text{variance of the product observations}}&&
\end{flalign}
\myequations{Intraclass correlation coefficient}

\begin{flalign}
    \overline{X}  \pm \text{CI} & = \frac{\sum\limits_{i=1}^{n} X_i}{n} \pm t_{1- \alpha / 2 \, \text{,} \, n-1} \times \frac{s}{\sqrt{n}}&&
\end{flalign}
\myequations{Average $\pm$ confidence interval}

\begin{flalign}
    \text{distance to target} & = \text{specification target} - \text{location statistic} = T - \overline{X}&&
\end{flalign}
\myequations{Distance to target}

\begin{flalign}
    s \pm \text{CI} & = \sqrt{\frac{\sum\limits_{i=1}^{n} \left(X_i - \overline{X}\right)^2}{n - 1}} \pm \left(\sqrt{\frac{(n - 1)\,s^{2}}{\chi_{\alpha / 2}^2}},\sqrt{\frac{(n - 1)\,s^{2}}{\chi_{1 - {\alpha / 2}}^2}}\right)&&
\end{flalign}
\myequations{Standard deviation $\pm$ confidence interval}

\begin{flalign}
    \text{distance to nearest specification} & = \frac{\text{nearest specification} - \text{location statistic}}{\text{dispersion statistic}}&& \\
                                             & = \frac{\text{min}\,(\overline{X} - \text{LSL}\,, \text{USL} - \overline{X})}{s}&&
\end{flalign}
\myequations{Distance to nearest specification}

The confidence intervals are:

\begin{flalign}
    \text{Lower bound} = \hat{P}_{pk} - z_{1 - \alpha / 2} \sqrt{\frac{1}{\left(\frac{\text{toler}}{2}\right)^2 n} + \frac{\left(\hat{P}_{pk}\right)^2}{2 \nu}}&&
\end{flalign}
\myequations{$\hat{P}_{pk}$ lower bound}

\begin{flalign}
    \text{Upper bound} = \hat{P}_{pk} + z_{1- \alpha / 2} \sqrt{\frac{1}{\left(\frac{\text{toler}}{2}\right)^2 n} + \frac{\left(\hat{P}_{pk}\right)^2}{2 \nu}}&&
\end{flalign}
\myequations{$\hat{P}_{pk}$ upper bound}

where:

\begin{flalign*}
    n                           & = \text{the number of observations in the sample}&& \\
    \nu                         & = \text{the degrees of freedom $= n - 1$}&& \\
    \chi_{\alpha \text{,}\nu}^2 & = \text{the $\alpha$ percentile of the chi-square distribution with $\nu$ degrees of freedom}&& \\
    \text{toler}                & = \text{multiplier of the standard deviation tolerance (Minitab\textregistered\ uses 6 as the default value)}&& \\
    z_{1 - \alpha / 2}          & = \text{the $1 - \alpha / 2$ percentile of the standard normal distribution}&& \\
\end{flalign*}

\begin{flalign}
A=\begin{pmatrix}
1 &0 &4 &-1 \\
0 &1 &-3 &1 \\
0 &0 &0 &1
\end{pmatrix}&&
\end{flalign}

\begin{flalign}
A = \begin{bmatrix}
1 &0 &4 &-1\\
0 &1 &-3 &1 \\
0 &0 &0 &1
\end{bmatrix}&&
\end{flalign}

\newpage
\section{Greek letters}\label{sec:greek_letters}

\subsection{Greek Unicode}\label{sec:greek_unicode}

Greek letters can be inserted in mathematics and text using Unicode characters. In Neovim, the sequence is `ControlKey + v uhhhh` where `hhhh` is the Unicode.

\begin{table}[H]
\begin{flushleft}
\begin{tabular}{lcccc}
\toprule
letter & code & miniscule & code & majuscule \\
\midrule
alpha   & 03b1 & α & 0391 & Α \\
beta    & 03b2 & β & 0392 & Β \\
gamma   & 03b3 & γ & 0393 & Γ \\
delta   & 03b4 & δ & 0394 & Δ \\
epsilon & 03b5 & ε & 0395 & Ε \\
zeta    & 03b6 & ζ & 0396 & Ζ \\
eta     & 03b7 & η & 0397 & Η \\
theta   & 03b8 & θ & 0398 & Θ \\
iota    & 03b9 & ι & 0399 & Ι \\
kappa   & 03ba & κ & 039a & Κ \\
lambda  & 03bb & λ & 039b & Λ \\
mu      & 03bc & μ & 039c & Μ \\
nu      & 03bd & ν & 039d & Ν \\
xi      & 03be & ξ & 039e & Ξ \\
omicron &      & o &      & O \\
pi      & 03c0 & π & 03ao & Π \\
rho     & 03c1 & ρ & 03a1 & Ρ \\
sigma   & 03c3 & σ & 03a3 & Σ \\
tau     & 03c4 & τ & 03a4 & Τ \\
upsilon & 03c5 & υ & 03a5 & Υ \\
phi     & 03c6 & φ & 03a6 & Φ \\
chi     & 03c7 & χ & 03a7 & Χ \\
psi     & 03c8 & ψ & 03a8 & Ψ \\
omega   & 03c9 & ω & 03a9 & Ω \\
\bottomrule
\end{tabular}
\label{tab:greek_unicode}
\caption{Greek Unicode}
\end{flushleft}
\end{table}

\newpage
\subsection{Greek mathematics}\label{sec:greek_mathematics}

Greek letters can be inserted in mathematics. These are naturally italicized.

\begin{table}[H]
\begin{flushleft}
\begin{tabular}{lcccc}
\toprule
letter & code & miniscule & code & majuscule \\
\midrule
alpha   & \$\textbackslash alpha\$   & $\alpha$   & \$A\$                      & $A$ \\
beta    & \$\textbackslash beta\$    & $\beta$    & \$B\$                      & $B$ \\
gamma   & \$\textbackslash gamma\$   & $\gamma$   & \$\textbackslash Gamma\$   & $\Gamma$ \\
delta   & \$\textbackslash delta\$   & $\delta$   & \$\textbackslash Delta\$   & $\Delta$ \\
epsilon & \$\textbackslash epsilon\$ & $\epsilon$ & \$E\$                      & $E$ \\
zeta    & \$\textbackslash zeta\$    & $\zeta$    & \$Z\$                      & $Z$ \\
eta     & \$\textbackslash eta\$     & $\eta$     & \$H\$                      & $H$ \\
theta   & \$\textbackslash theta\$   & $\theta$   & \$\textbackslash Theta\$   & $\Theta$ \\
iota    & \$\textbackslash iota\$    & $\iota$    & \$I\$                      & $I$ \\
kappa   & \$\textbackslash kappa\$   & $\kappa$   & \$K\$                      & $K$ \\
lambda  & \$\textbackslash lambda\$  & $\lambda$  & \$\textbackslash Lambda\$  & $\Lambda$ \\
mu      & \$\textbackslash mu\$      & $\mu$      & \$M\$                      & $M$ \\
nu      & \$\textbackslash nu\$      & $\nu$      & \$N\$                      & $N$ \\
xi      & \$\textbackslash xi\$      & $\xi$      & \$\textbackslash Xi\$      & $\Xi$ \\
omicron & \$o\$                      & $o$        & \$O\$                      & $O$ \\
pi      & \$\textbackslash pi\$      & $\pi$      & \$\textbackslash Pi\$      & $\Pi$ \\
rho     & \$\textbackslash rho\$     & $\rho$     & \$P\$                      & $P$ \\
sigma   & \$\textbackslash sigma\$   & $\sigma$   & \$\textbackslash Sigma\$   & $\Sigma$ \\
tau     & \$\textbackslash tau\$     & $\tau$     & \$T\$                      & $T$ \\
upsilon & \$\textbackslash upsilon\$ & $\upsilon$ & \$\textbackslash Upsilon\$ & $\Upsilon$ \\
phi     & \$\textbackslash phi\$     & $\phi$     & \$\textbackslash Phi\$     & $\Phi$ \\
chi     & \$\textbackslash chi\$     & $\chi$     & \$X\$                      & $X$ \\
psi     & \$\textbackslash psi\$     & $\psi$     & \$\textbackslash Psi\$     & $\Psi$ \\
omega   & \$\textbackslash omega\$   & $\omega$   & \$\textbackslash Omega\$   & $\Omega$ \\
\bottomrule
\end{tabular}
\label{tab:greek_mathematics}
\caption{Greek mathematics}
\end{flushleft}
\end{table}

\newpage
\subsection{Greek text}\label{sec:greek_text}

Greek letters can be inserted as text.

\begin{table}[H]
\begin{flushleft}
\begin{tabular}{lcccc}
\toprule
letter & code & miniscule & code & majuscule \\
\midrule
alpha    & \textbackslash textalpha   & \textalpha   & \textbackslash textAlpha   & \textAlpha \\
beta     & \textbackslash textbeta    & \textbeta    & \textbackslash textBeta    & \textBeta \\
gamma    & \textbackslash textgamma   & \textgamma   & \textbackslash textGamma   & \textGamma \\
delta    & \textbackslash textdelta   & \textdelta   & \textbackslash textDelta   & \textDelta \\
epsilon  & \textbackslash textepsilon & \textepsilon & \textbackslash textEpsilon & \textEpsilon \\
zeta     & \textbackslash textzeta    & \textzeta    & \textbackslash textZeta    & \textZeta \\
eta      & \textbackslash texteta     & \texteta     & \textbackslash textEta     & \textEta \\
theta    & \textbackslash texttheta   & \texttheta   & \textbackslash textTheta   & \textTheta \\
iota     & \textbackslash textiota    & \textiota    & \textbackslash textIota    & \textIota \\
kappa    & \textbackslash textkappa   & \textkappa   & \textbackslash textKappa   & \textKappa \\
lambda   & \textbackslash textlambda  & \textlambda  & \textbackslash textLambda  & \textLambda \\
mu       & \textbackslash textmu      & \textmu      & \textbackslash textMu      & \textMu \\
nu       & \textbackslash textnu      & \textnu      & \textbackslash textNu      & \textNu \\
xi       & \textbackslash textxi      & \textxi      & \textbackslash textXi      & \textXi \\
omicrdon & o                          & o            & O                          & O \\
pi       & \textbackslash textpi      & \textpi      & \textbackslash textPi      & \textPi \\
rho      & \textbackslash textrho     & \textrho     & \textbackslash textRho     & \textRho \\
sigma    & \textbackslash textsigma   & \textsigma   & \textbackslash textSigma   & \textSigma \\
tau    & \textbackslash texttau   & \texttau   & \textbackslash textTau   & \textTau \\
upsilon    & \textbackslash textupsilon   & \textupsilon   & \textbackslash textUpsilon   & \textUpsilon \\
phi    & \textbackslash textphi   & \textphi   & \textbackslash textPhi   & \textPhi \\
chi    & \textbackslash textchi   & \textchi   & \textbackslash textChi   & \textChi \\
psi    & \textbackslash textpsi   & \textpsi   & \textbackslash textPsi   & \textPsi \\
omega    & \textbackslash textomega   & \textomega   & \textbackslash textOmega   & \textOmega \\
\bottomrule
\end{tabular}
\label{tab:greek_text}
\caption{Greek text}
\end{flushleft}
\end{table}

\newpage
\section{Theorems}\label{sec:theorems}

\begin{theorem}[Fundamental Theorem of Calculus]
\label{thm:ftc}
If $f$ is continuous on the interval $[a,b]$, then \[\frac{d}{dx}\int_a^x f(t)dx=f(x)\] for all $a\le x\le b$.
\end{theorem}
\mytheorems{Fundamental theorem of calculus}

\section{Definitions}\label{sec:definitions}

\begin{definition}[Definition of limit]
\label{def:limit}
Let $f$ be a function defined on an open interval containing $x=a$, but perhaps not at $x=a$.  We say $\displaystyle\lim_{x\to a}f(x)=L$ if for any $\epsilon>0$ there exists $\delta>0$ such that $|f(x)-L|<\epsilon$ whenever $0<|x-a|<\delta$.
\end{definition}
\mydefinitions{Definition of limit}

\begin{definition}
\label{def:derivative}
The \textit{derivative} of $f$ is defined to be $\displaystyle f^\prime(x)=\lim_{h\to 0}\frac{f(x+h)-f(x)}{h}$, if this limit exists.
\end{definition}
\mydefinitions{Derivative of f}

\section{Corollaries}\label{sec:corollaries}
\begin{corollary}
If $f$ is continuous on the interval $[a,b]$, then \[\int_a^b f(x)dx=F(b)-F(a)\] where $F$ is any antiderivative of $f$.
\end{corollary}
\mycorollaries{f is continuous}

\newpage
\section{Lemmas}\label{sec:lemmas}
\begin{lemma}
$x+y=z$
\end{lemma}
\mylemmas{Lem 1}

\begin{lemma}
$z-r+t=3$
\end{lemma}
\mylemmas{Lem 2}

\begin{lemma}
$q=r$
\end{lemma}
\mylemmas{Lem 3}

\begin{lemma}
$x+y=z$
\end{lemma}
\mylemmas{Lem 4}

\begin{lemma}
$z-r+t=3$
\end{lemma}
\mylemmas{Lem 5}

\begin{lemma}
$q=r$
\end{lemma}
\mylemmas{Lem 6}

\newpage
\section{Propositions}\label{sec:propositions}
\begin{proposition}
$x+y=z$
\end{proposition}
\mypropositions{Prop 1}

\begin{proposition}
$\alpha=1-x$
\end{proposition}
\mypropositions{Prop 2}

\newpage
\section{Citations}\label{sec:citations}

Here is a single citation without prenote and postnote. There is a definitive book on MSA (\cite{wheeler2006}).

Here is a single citation with prenote only. There is a definitive book on SPC (\cite[see][]{wheeler2006}).

Here is a single citation with postnote only. $C_\text{pm}$ is a complex subject (\cite[][pages 1-10]{chan1988}).

\newpage
\listoffigures

\newpage
\listoftables

\newpage
\listofmyequations

\newpage
\listofmytheorems

\newpage
\listofmydefinitions

\newpage
\listofmycorollaries

\newpage
\listofmylemmas

\newpage
\listofmypropositions

\newpage
\section{Acknowledgements}\label{sec:acknowledgements}

The author wishes to thank ...

\section{Supporting information}\label{sec:supporting_information}

Add files, data, and other supporting information, and where it can be obtained.

\newpage
\section{References}\label{sec:references}
\nocite{*}
\printbibliography[heading=none]

\end{document}