### 2022-07-21 pilon_gilles_glossary.tex
Create pilon_gilles_glossary.tex

\newglossaryentry{process_capability}
{
    name={process capability},
    description={The ability of a process to meet a performance standard}
}
\newglossaryentry{process_capability_index}
{
    name={process capability index},
    description={The ratio of a specification measure and a dispersion measure},
    plural={process capability indices}
}
\newdualentry{msa}{MSA}{Measurement System Analysis}{Methods to evaluate measurements systems, to understand the causes of measurement variation}
\newglossaryentry{test_retest_error}
{
    name={test-retest error},
    description={The standard deviation of making repeated measurements of the same characteristic on the same parts with the same measuring instrument}
}
\newglossaryentry{probable_error}
{
    name={probable error},
    description={TBD}
}
\newdualentry{icc}{ICC}{intraclass correlation coefficient}{Describes how strongly units in the same group resemble each other}
\newglossaryentry{statistical_control}
{
    name={statistical control},
    description={A process is said to be in a state of statistical control if the variation for both charts is stable and predictable. This means that no points fail various control chart rules and no systematic patterns of variation are present.}
}
\newglossaryentry{distance_to_nearest_specification}
{
    name={distance to nearest specification (sigma level)},
    description={The difference between the location statistic and the closest specification limit, divided by the dispersion statistic (e.g. the standard deviation)}
}
\newglossaryentry{hypothesis_test}
{
    name={hypothesis test},
    description={A method for proving that an observed difference was not due to random chance},
    plural={hypothesis tests}
}
\newglossaryentry{dot_plot}
{
    name={dot plot},
    description={A chart in which dots are used to depict the quantitative values (e.g. counts) associated with categorical variables},
    plural={dot plots}
}
\newglossaryentry{normal_distribution}
{
    name={normal distribution},
    description={A type of continuous probability distribution for a real-valued random variable},
    plural={normal distributions}
}
\newglossaryentry{confidence_interval}
{
    name={confidence interval},
    description={It is an interval calculated from values in a sample. The interval is constructed for a specified probability of containing the unknown true value of a population parameter.
    },
    plural={confidence intervals}
}