### 2025-06-13 curriculum_vitae.tex
Create curriculum_vitae.tex

\documentclass[10pt, letterpaper, twoside]{article} % Using article class for simplicity
\usepackage[utf8]{inputenc}     % For correct handling of special characters
\usepackage[sfdefault]{noto}
\usepackage[T1]{fontenc}        % For proper font encoding
\usepackage[nohyper]{ragged2e}  % For \RaggedRight to left-align content
\usepackage{geometry}           % For easy margin setting
\usepackage{url}                % For formatting URLs
\usepackage{titlesec}           % For controlling section spacing
\usepackage{enumitem}           % For controlling list indentation

% --- Custom Section Formatting ---
% \titleformat{<command>}[<shape>]{<format>}{<label>}{<sep>}{<before-code>}[<after-code>]
\titleformat{\section}[block]{\bfseries\large}{\thesection}{1em}{}

% --- Geometry Settings ---
\geometry{margin=0.5in} % Set all margins to 1 inch

% --- Page Style (No Page Numbers) ---
\pagestyle{empty}

% --- Reduce Section Spacing ---
% \titlespacing{<command>}{<left>}{<before-skip>}{<after-skip>}
\titlespacing{\section}{0pt}{12pt}{5pt}
\titlespacing{\subsection}{0pt}{8pt}{3pt}

% --- Custom Em Dash Macro (if you defined it) ---
\newcommand{\spacedemdash}{\,---\,} % Example: if you want thin spaces

% --- Global Itemize Indentation & Spacing Control (MODIFIED) ---
\setlist[itemize]{
    leftmargin=*,      % Keeps text flush with left margin
    topsep=0pt,        % Reduces space before the first item
    partopsep=0pt,     % Reduces additional space if list starts new paragraph
    itemsep=0pt,       % Reduces space between individual items
    parsep=0pt         % Reduces space between paragraphs *within* an item (if any)
}

% --- Discourage hyphenation
\hyphenpenalty=10000   % Discourage hyphenation of words
\exhyphenpenalty=10000 % Discourage hyphenation of words already containing hyphens

\begin{document}
% --- Apply \RaggedRight to the entire document body (excluding centered parts) ---
\RaggedRight

% --- Contact Information ---
\begin{center}
    \textbf{\large FIRSTNAME LASTNAME} \\
    city province \\
    telephone \\
    email \\
    \url{https://github.com/yourid} \\
    \url{https://www.linkedin.com/in/yourid/}
\end{center}

% --- Professional Title ---
% \RaggedRight ensures the content is left-aligned from here onwards.
% Place it before the content you want to affect.
% \centerline{\textbf{Senior Machine Learning Engineer}}

\section*{SUMMARY}
insert text

\section*{EXPERIENCE}

\noindent\textbf{company \textbar ~city province \textbar ~title} \hfill \textbf{yyyy-mm\spacedemdash yyyy-mm} \\
brief job description
\begin{itemize}
    \item achievement
    \item achievement
\end{itemize}
\vspace{8pt}

\noindent\textbf{company \textbar ~city province \textbar ~title} \hfill \textbf{yyyy-mm\spacedemdash yyyy-mm} \\
brief job description
\begin{itemize}
    \item achievement
    \item achievement
\end{itemize}
\vspace{8pt}

\section*{SKILLS}
skill, skill, ...

\section*{EDUCATION}
degree, major, university, city province

\end{document}